---
title: "Data_analysis"
output: html_document
---

#loading required libraries
```{r}
library(tidyverse)
library(lubridate)
library(showtext)
library(extrafont)
library(ggtext)
library(ggthemes)
library(patchwork)
library(gt)
library(gtExtras)
#font_import()
#loadfonts(device = "win")


```

#setting ubuntu theme
```{r}
#setting theme
ubuntu_theme<-function()
{
  #setting font
Font<-"Comic Sans MS"

theme_economist() %+replace%
  
  theme(
    ########grid elements########
    panel.grid.major = element_blank(),    #strip major gridlines
    panel.grid.minor = element_blank(),    #strip minor gridlines
     axis.ticks = element_blank(),          #strip axis ticks
    axis.line = element_blank(),
    ######text elements###########
    plot.title = element_markdown(             #title
                   family = Font,            #set font family
                   size = 13,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = 0,                #left align
                   vjust = 2),               #raise slightly
      
      plot.subtitle = element_markdown(          #subtitle
                   family = Font,            #font family
                   size = 12,hjust = 0),               #font size
      
      plot.caption = element_text(           #caption
                   family = Font,            #font family
                   size = 8.7,
                   hjust = 0
                   ),               #right align
      
      axis.title = element_text( family = Font,size = 9),               #font size
      
      axis.text = element_text( family = Font,size = 9),               #font size
    
      legend.text = element_text( family = Font,size = 9),
      legend.title = element_text( family = Font,size = 9),
      legend.position = "right",
      plot.background = element_rect(fill = "white", color = NA)

    
    ###########plot#################
    
  )
}
```

#loading data
```{r}
df<-read_csv("./www/report_data.csv")
```

#exploring data
```{r}
head(df)
```
#checking for NAs per column
```{r}
colSums(is.na(df))
```

#data wrangling
```{r}
df <-df %>% mutate(
  Team=case_when(Name=="Barasa Murefu"~"Team Delta",
                 Name=="Kinuthia Brian"~"Team Phi",
                 Name=="Shikorire Tom"~"Team Phi",
                 Name=="Anyango Ruth"~"Team Phi",
                 Name=="Mwangi Brian"~"Team Phi",
                 Name=="Tanui Cornelius"~"Team Phi",
                 Name=="Mbutura Alois"~"Team Kappa",
                 Name=="Kitetu Martin"~"Team Kappa",
                 Name=="Okode Adams"~"Team Kappa",
                 Name=="Irungu Joyce"~"Team Chi",
                 Name=="Odhiambo Elmah"~"Team Chi",
                 Name=="Samo June"~"Team Chi",
                 Name=="Cherunya Pauline"~"Team Chi",
                 Name=="Odero Faith"~"Team Chi",
                 Name=="Musivo Winnie"~"Team Gamma",
                 Name=="Gichuhi Ruth"~"Team Gamma",
                 Name=="Muchiri James"~"Team Gamma",
                 Name=="Maina Morris"~"Team Gamma",
                 Name=="Kamau Clara"~"Team Gamma",
                 Name=="Muthuri Cynthia"~"Team Gamma"
                 ),
  Rate=case_when(Name=="Barasa Murefu"~"734",
                 Name=="Kinuthia Brian"~"490",
                 Name=="Shikorire Tom"~"415",
                 Name=="Anyango Ruth"~"282",
                 Name=="Mwangi Brian"~"282",
                 Name=="Tanui Cornelius"~"282",
                 Name=="Mbutura Alois"~"360",
                 Name=="Kitetu Martin"~"445",
                 Name=="Okode Adams"~"360",
                 Name=="Irungu Joyce"~"415",
                 Name=="Odhiambo Elmah"~"282",
                 Name=="Samo June"~"445",
                 Name=="Cherunya Pauline"~"445",
                 Name=="Odero Faith"~"282",
                 Name=="Musivo Winnie"~"415",
                 Name=="Gichuhi Ruth"~"445",
                 Name=="Muchiri James"~"282",
                 Name=="Maina Morris"~"282",
                 Name=="Kamau Clara"~"415",
                 Name=="Muthuri Cynthia"~"360"
                 ),
  Sector=case_when(Project=="Ace Taf" ~"Energy",
                 Project=="AECF REACT" ~"Energy",
                 Project=="AECF REACT Somalia"~"Energy",
                 Project=="AECF Refresh"~"Energy",
                 Project== "AECF Tanzania" ~"Energy",
                 Project=="AfR"~"Climate Change",
                 Project=="AMDA" ~"Energy",
                 Project=="Burundi MTF" ~"Energy",
                 Project=="Burundi MTF Expansion" ~"Energy",
                 Project=="CCA" ~"Energy",
                 Project=="CDC" ~"Energy",
                 Project=="CLASP Field Monitoring" ~"Energy",
                 Project=="CLASP QV"  ~"Energy",
                 Project=="CLI 1"~"Climate Change",
                 Project=="CLI 2"~"Climate Change",
                 Project=="CTCN Urban  Briquette"~"Energy",
                 Project=="Data Collection for Emission Quantification in Shipping(GIZ)"~"Climate Change",
                 Project=="EED Admin Qtr 1/21" ~"Admin",
                 Project=="EED Admin Qtr 2/21" ~"Admin",
                 Project=="EED Admin Qtr 3/21" ~"Admin",
                 Project=="EED Admin Qtr 4/21" ~"Admin",
                 Project=="EED BD Qtr 1/21"  ~"Business Development",
                 Project=="EED BD Qtr 2/21"  ~"Business Development",
                 Project=="EED BD Qtr 3/21"  ~"Business Development",
                 Project=="EED BD Qtr 4/21"  ~"Business Development",
                 Project=="EED BD Qtr 4/20"  ~"Business Development",
                 Project=="Energy Access Explorer" ~"Energy",
                 Project=="EPRA GMG"  ~"Energy",
                 Project=="Green Finance- FSD MoC"  ~"Climate Change",
                 Project=="HavGas"   ~"Energy",
                 Project=="Hygiene Market Analysis" ~"Water",
                 Project=="IFC Uganda"   ~"Energy",
                 Project=="IMEP Baseline Survey GMG"   ~"Energy",
                 Project=="Kaiote Admin Qtr 1/21"  ~"Admin",
                 Project=="Kaiote Admin Qtr2 /21"  ~"Admin",
                 Project=="Kaiote Admin Qtr 3/21"  ~"Admin",
                 Project=="Kaiote Admin Qtr 4/21"  ~"Admin",
                 Project=="Kaiote BD Qtr 1/21"  ~"Business Development",
                 Project=="Kaiote BD Qtr 2/21"  ~"Business Development",
                 Project=="Kaiote BD Qtr 4/21"  ~"Business Development",
                 Project=="KawiSafi Technical Assistance Facility (TAF)"  ~"Energy",
                 Project=="Kenya Rapid MWA" ~"Water",
                 Project=="KfW"   ~"Energy",
                 Project=="Leave 2021"  ~"Leave",
                 Project=="LMCP Phase 1 AFDB"   ~"Energy",
                 Project=="MOU development Karibu Homes Ltd and MAVWASCO"   ~"Water",
                 Project=="MWA 2"  ~"Water",
                 Project=="Nakuru Energy Plan"  ~"Energy",
                 Project=="NCI Report Review"   ~"Climate Change",
                 Project=="Power to X"   ~"Energy",
                 Project=="PPA Watch"   ~"Energy",
                 Project=="RE4R Implementation"   ~"Energy",
                 Project=="RESI"   ~"Energy",
                 Project=="Retreat"  ~"Admin",
                 Project=="Roadmap to E-mobility" ~"Climate Change",
                 Project=="TASC Improved cookstove and clean water POA"   ~"Energy",
                 Project=="UNICEF KOICA"     ~"Water",
                 Project=="US NIH" ~"Energy",
                 Project=="WB Ethiopia"    ~"Energy",
                 Project=="WRI spatial data collection" ~"Energy",
                 ),
   Lines=case_when(Sector=="Energy"~"Consulting",
                 Sector=="Climate Change"~"Consulting",
                 Sector=="Water" ~"Consulting",
                 Sector=="Admin"   ~"Admin",
                 Sector=="Business Development"   ~"Business Development",
                 Sector=="Leave"    ~"Leave"
                 ),
  Rate=as.numeric(Rate),
  Cost=round(Rate*Hrs/8,0),
  Date=dmy(Date),
  Quarter=quarter(Date)
  )
```

#key lines data for plotting
```{r}
lines_year<-df %>% 
  group_by(Lines) %>%
  summarise(total_hrs=round(sum(Hrs),0),
            total_cost=sum(Cost)
           ) %>%
  mutate(
     hrs_perc=round(total_hrs/sum(total_hrs)*100,0)
  )


lines_year[2,1]<-"BD"
lines_year
```


```{r}
f1<-"Comic Sans MS"
f2<-"Palentino Linotype"
```



#plotting lines per year
```{r}
# Compute percentages
lines_year$fraction <- lines_year$hrs_perc / sum(lines_year$hrs_perc)

# Compute the cumulative percentages (top of each rectangle)
lines_year$ymax <- cumsum(lines_year$fraction)

# Compute the bottom of each rectangle
lines_year$ymin <- c(0, head(lines_year$ymax, n=-1))

# Compute label position
lines_year$labelPosition <- (lines_year$ymax + lines_year$ymin) / 2

# Compute a good label
lines_year$label <- paste0(lines_year$Lines, "\n", lines_year$hrs_perc,"%")

##ggplot
p1<-ggplot(lines_year, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Lines)) +
  geom_rect() +
  geom_text( x=4.6, aes(y=labelPosition, label=label,color=Lines), size=3.2,family=f1) +
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107")) +
  scale_color_manual(values = c("#008253","#393d83","#fd7e14","#ffc107")) +
  coord_polar(theta="y") +
  xlim(c(1, 4)) +
  theme_void() +
  theme(legend.position = "none")
ggsave("./www/lines_year.png",
       dpi = 320,
      width = 10,
      height = 7)
p1
```

#sector lines data for plotting
```{r}
sector_year<-df %>% 
  group_by(Sector) %>%
  summarise(total_hrs=round(sum(Hrs),0),
            total_cost=sum(Cost)
           ) %>%
  mutate(
     hrs_perc=round(total_hrs/sum(total_hrs)*100,1)
  )
sector_year

sct<-c("Climate Change","Energy","Water")

eed_sct<-sector_year %>% filter(Sector %in% sct)

eed_sct
```

#stacked plot for sector lines
```{r}
p2<-ggplot(eed_sct,aes(x="",y=hrs_perc,fill=Sector))+
  geom_bar(position = "stack",stat = "identity",width = 0.2)+
  geom_text(aes(label=paste0(Sector,"\n",hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=2.5)+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14"))+
  # geom_curve(aes(x = 40, y = 3, xend = 0, yend = -6),
  #            size = 0.5, color = "black",curvature = -0.2,
  #            arrow = arrow(length = unit(0.02, "npc")))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    axis.title = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "None",
    legend.title = element_blank()
  )
p2
  
```

#layout
```{r}
p3<-(p1|p2)
ggsave("./www/lines_output2.png",plot = p3, dpi = 320, height = 8, width = 11)
```

#yearly project costs
```{r}
project_year<-df %>% 
  group_by(Project) %>%
  summarise(total_cost=round(sum(Cost),0)) %>%
  arrange(desc(total_cost)) 

#selecting rows with projects only
project_year<-project_year[-c(1,4,7,9,12:16,19,20,24,30,43,49,50,53),]
project_year
```

#yearly project cost per sector
```{r}
sct_yr<-df %>% 
  group_by(Project,Sector) %>%
  summarise(total_cost=round(sum(Cost),0),
            total_hrs=round(sum(Hrs),0)) %>%
  arrange(desc(total_cost)) 
sct_yr<-sct_yr[-c(1,4,7,9,12:16,19,20,24,30,43,49,50,53),]
sct_energy<-sct_yr %>% filter(Sector=="Energy")
sct_water<-sct_yr %>% filter(Sector=="Water")
sct_climate<-sct_yr %>% filter(Sector=="Climate Change") %>% select(-Sector)
sct_yr

```

#table per sector- water
```{r}
write.csv(sct_water,"./www/sct_water2.csv")
sct_water2<-read_csv("./www/sct_water2.csv")
sct_watertb<-sct_water2 %>%
  rename(
    'Total cost'=total_cost,
    'Total Hours'=total_hrs
  ) %>%
  gt() %>%
  fmt_currency(
    columns = 'Total cost',
    currency = "USD",
    sep_mark = ",",
    decimals = NULL
  ) %>%
  gt_color_rows('Total cost':'Total Hours',palette = "ggsci::blue_material" ) %>%

   #rename columns
  cols_label(
    'Total cost'="Total cost ($)"
  ) %>%
  gt_theme_guardian()

sct_watertb %>% gtsave("./www/sct_water2.png")
```


#table per sector- climate
```{r}
write.csv(sct_climate,"./www/sct_climate2.csv")
sct_climate2<-read_csv("./www/sct_climate2.csv")
sct_climatetb<-sct_climate2 %>%
  rename(
    'Total cost'=total_cost,
    'Total Hours'=total_hrs
  ) %>%
  gt() %>%
  fmt_currency(
    columns = 'Total cost',
    currency = "USD",
    sep_mark = ",",
    decimals = NULL
  ) %>%
  gt_color_rows('Total cost':'Total Hours',palette = "ggsci::blue_material" ) %>%

   #rename columns
  cols_label(
    'Total cost'="Total cost ($)"
  ) %>%
  # tab_header(
  #   title = "CLIMATE PROJECTS",
  #   #use markdown syntax with md()
  #   subtitle = md("**Summary**")
  # ) %>% 
  #  #customizing the title
  # tab_style(
  #   #select object to modify
  #   locations = cells_title(groups ="title"),
  #   #specifying text style
  #   style = list(
  #     cell_text(
  #       font=f2,
  #       size = "xx-large",
  #       color = "#393d83",
  #       align = "center"
  #     )
  #   )
  # ) %>% 
  #  tab_style(
  #   #select object to modify
  #   locations = cells_title(groups ="subtitle"),
  #   #specifying text style
  #   style = list(
  #     cell_text(
  #       font=f2,
  #       size = "x-large",
  #       color = "#393d83",
  #       align = "center"
  #     )
  #   )
  # ) %>% 
  gt_theme_guardian()

sct_climatetb %>% gtsave("./www/sct_climate2.png")
```

#table per sector- energy
```{r}
write.csv(sct_energy,"./www/sct_energy2.csv")
sct_energy2<-read_csv("./www/sct_energy2.csv")
sct_energytb<-sct_energy2 %>%
  rename(
    'Total cost'=total_cost,
    'Total Hours'=total_hrs
  ) %>%
  gt() %>%
   fmt_currency(
    columns = 'Total cost',
    currency = "USD",
    sep_mark = ",",
    decimals = NULL
  ) %>%
  gt_color_rows('Total cost':'Total Hours',palette = "ggsci::blue_material" ) %>%

   #rename columns
  cols_label(
    'Total cost'="Total cost ($)"
  ) %>%
  tab_style(
    #select object to modify
    locations = cells_title(groups ="cell"),
    #specifying text style
    style = list(
      cell_text(
        size = "xx-large"
      )
    )
  ) %>%
  # tab_header(
  #   title = "ENERGY PROJECTS",
  #   #use markdown syntax with md()
  #   subtitle = md("**Summary**")
  # ) %>% 
  #  #customizing the title
  # tab_style(
  #   #select object to modify
  #   locations = cells_title(groups ="title"),
  #   #specifying text style
  #   style = list(
  #     cell_text(
  #       font=f2,
  #       size = "xx-large",
  #       color = "#393d83",
  #       align = "center"
  #     )
  #   )
  # ) %>% 
  #  tab_style(
  #   #select object to modify
  #   locations = cells_title(groups ="subtitle"),
  #   #specifying text style
  #   style = list(
  #     cell_text(
  #       font=f2,
  #       size = "x-large",
  #       color = "#393d83",
  #       align = "center"
  #     )
  #   )
  # ) %>% 
  gt_theme_guardian()

sct_energytb
sct_energytb %>% gtsave("./www/sct_energy2.png")
```

#saving plot
```{r}

```

#team yearly metrics
```{r}
team_lines<-df %>% 
  group_by(Team,Lines) %>%
  summarise(total_hrs=round(sum(Hrs),0),
            total_cost=sum(Cost)
           ) %>%
  mutate(
     hrs_perc=round(total_hrs/sum(total_hrs)*100,1)
  )
team_lines

```

#plotting team yearly metrics
```{r}
ggplot(team_lines,aes(x=Team,y=hrs_perc,fill=Lines))+
  geom_bar(position = "stack",stat = "identity")+
  geom_text(aes(label=paste0(hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=3)+
  labs(y="Percentage")+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )
#ggsave("./www/team_lines2.png")
```

#personal yearly metrics
```{r}
personal_lines<-df %>% 
  group_by(Name,Lines) %>%
  summarise(total_hrs=round(sum(Hrs),0),
            total_cost=sum(Cost)
           ) %>%
  mutate(
     hrs_perc=round(total_hrs/sum(total_hrs)*100,1)
  )


lft<-c("Kamau Clara","Musivo Winnie","Odero Faith","Tanui Cornelius")
personal_lines<-personal_lines %>% filter(!Name %in% lft)
personal_lines2<-personal_lines %>% 
  mutate(
    Team=case_when(Name=="Barasa Murefu"~"Team Delta",
                 Name=="Kinuthia Brian"~"Team Phi",
                 Name=="Shikorire Tom"~"Team Phi",
                 Name=="Anyango Ruth"~"Team Phi",
                 Name=="Mwangi Brian"~"Team Phi",
                 Name=="Mbutura Alois"~"Team Kappa",
                 Name=="Kitetu Martin"~"Team Kappa",
                 Name=="Okode Adams"~"Team Kappa",
                 Name=="Irungu Joyce"~"Team Chi",
                 Name=="Odhiambo Elmah"~"Team Chi",
                 Name=="Samo June"~"Team Chi",
                 Name=="Cherunya Pauline"~"Team Chi",
                 Name=="Gichuhi Ruth"~"Team Gamma",
                 Name=="Muchiri James"~"Team Gamma",
                 Name=="Maina Morris"~"Team Gamma",
                 Name=="Muthuri Cynthia"~"Team Gamma"
                 )
  ) %>% 
  select(-c(3,4))
personal_lines2
personal_lines3<-personal_lines2 %>% spread(Lines,hrs_perc) %>% select(-Team)

write.csv(personal_lines3,"./www/lines_table.csv",row.names = FALSE)
personal_lines4<-read_csv("./www/lines_table.csv")
```

#general table for time spent
```{r}
pl4tb<-personal_lines4 %>%
  gt() %>%
  fmt_percent(
    columns = 'Admin':'Leave',
    scale_values = FALSE,
    decimals = 1
  )%>%
  gt_color_rows('Admin':'Leave',palette = "ggsci::blue_material" ) %>%
  
  gt_theme_guardian()

pl4tb %>% gtsave("./www/lines_table2.png")
```


#plot personal_lines team phi
```{r}
personal_lines2 %>% 
  filter(Team=="Team Phi")%>%
  ggplot(aes(x=Name,y=hrs_perc,fill=Lines))+
  geom_bar(position = "stack",stat = "identity")+
  geom_text(aes(label=paste0(hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=3)+
  labs(title ="Team Phi",y="Percentage",x="")+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )
ggsave("./www/tphi_yearly2.png")
```

#plot personal_lines team delta
```{r}
personal_lines2 %>% 
  filter(Team=="Team Delta")%>%
  ggplot(aes(x=Name,y=hrs_perc,fill=Lines))+
  geom_bar(position = "stack",stat = "identity")+
  geom_text(aes(label=paste0(hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=3)+
  labs(title ="Team Delta",y="Percentage",x="")+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )
ggsave("./www/tdelta_yearly2.png")

```

#plot personal_lines team chi
```{r}
personal_lines2 %>% 
  filter(Team=="Team Chi")%>%
  ggplot(aes(x=Name,y=hrs_perc,fill=Lines))+
  geom_bar(position = "stack",stat = "identity")+
  geom_text(aes(label=paste0(hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=3)+
  labs(title ="Team Chi",y="Percentage",x="")+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )
ggsave("./www/tchi_yearly2.png")
```


#plot personal_lines team Gamma
```{r}
personal_lines2 %>% 
  filter(Team=="Team Gamma")%>%
  ggplot(aes(x=Name,y=hrs_perc,fill=Lines))+
  geom_bar(position = "stack",stat = "identity")+
  geom_text(aes(label=paste0(hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=3)+
  labs(title ="Team Gamma",y="Percentage",x="")+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )
ggsave("./www/tgamma_yearly2.png")
```

#plot personal_lines team kappa
```{r}
personal_lines2 %>% 
  filter(Team=="Team Kappa")%>%
  ggplot(aes(x=Name,y=hrs_perc,fill=Lines))+
  geom_bar(position = "stack",stat = "identity")+
  geom_text(aes(label=paste0(hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=3)+
  labs(title ="Team Kappa",y="Percentage",x="")+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    legend.position ="bottom",
    legend.title = element_blank()
  )
ggsave("./www/tkappa_yearly2.png")
```

#time spent on each project per team
```{r}
tproject<- df %>% 
  filter(Lines=="Consulting") %>%
  group_by(Project,Team) %>%
  summarise(total_hrs=round(sum(Hrs),0),
            total_cost=sum(Cost)
           ) 
  
tproject
```
#team gamma
```{r}
tproject %>% 
  filter(Team=="Team Gamma") %>%
  ggplot(aes(x=reorder(Project,total_hrs),y=total_hrs))+
  geom_col(fill="#008253")+
  labs(title ="Team Gamma",y="Total hours",x="")+
  geom_text(aes(label=paste0(total_hrs)),position=position_dodge(width=1.2),color="black",
            hjust=0,vjust=0.5,family=f1,size=2.8)+
  coord_flip()+
  ubuntu_theme()
ggsave("./www/tproject_gamma.png")
```

#team phi
```{r}
tproject %>% 
  filter(Team=="Team Phi") %>%
  ggplot(aes(x=reorder(Project,total_hrs),y=total_hrs))+
  geom_col(fill="#393d83")+
  labs(title ="Team Phi",y="Total hours",x="")+
  geom_text(aes(label=paste0(total_hrs)),position=position_dodge(width=1.2),color="black",
            hjust=0,vjust=0.5,family=f1,size=2.6)+
  coord_flip()+
  ubuntu_theme()
ggsave("./www/tproject_phi.png")
```

#team chi
```{r}
tproject %>% 
  filter(Team=="Team Chi") %>%
  ggplot(aes(x=reorder(Project,total_hrs),y=total_hrs))+
  geom_col(fill="#fd7e14")+
  labs(title ="Team Chi",y="Total hours",x="")+
  geom_text(aes(label=paste0(total_hrs)),position=position_dodge(width=1.2),color="black",
            hjust=0.6,vjust=0.5,family=f1,size=2.4)+
  coord_flip()+
  ubuntu_theme()
ggsave("./www/tproject_chi.png")
```

#team kappa
```{r}
tproject %>% 
  filter(Team=="Team Kappa") %>%
  ggplot(aes(x=reorder(Project,total_hrs),y=total_hrs))+
  geom_col(fill="#ffc107")+
  labs(title ="Team Kappa",y="Total hours",x="")+
  geom_text(aes(label=paste0(total_hrs)),position=position_dodge(width=1.2),color="black",
            hjust=0.6,vjust=0.5,family=f1,size=2.4)+
  coord_flip()+
  ubuntu_theme()
ggsave("./www/tproject_kappa.png")
```

#Team Delta
```{r}
tproject %>% 
  filter(Team=="Team Delta") %>%
  ggplot(aes(x=reorder(Project,total_hrs),y=total_hrs))+
  geom_col(fill="#2596be")+
  labs(title ="Team Delta",y="Total hours",x="")+
  geom_text(aes(label=paste0(total_hrs)),position=position_dodge(width=1.2),color="black",
            hjust=0,vjust=0.5,family=f1,size=2.4)+
  coord_flip()+
  ubuntu_theme()
ggsave("./www/tproject_delta.png")
```

#time series analysis by lines
```{r}
library(zoo)
p_pattern<-df %>% 
  mutate(
    month=as.yearmon(Date),
  ) %>%
  group_by(Lines,month) %>%
  summarise(total_hrs=round(sum(Hrs),0),
            total_cost=sum(Cost)
           ) 
 

p_pattern$month<-as.Date.yearmon(p_pattern$month)
p_pattern
```

#time series analysis by sector
```{r}
s_pattern<-df %>% 
  filter(Lines=="Consulting") %>%
  mutate(
    month=as.yearmon(Date),
  ) %>%
  group_by(Sector,month) %>%
  summarise(total_hrs=round(sum(Hrs),0),
            total_cost=sum(Cost)
           ) 
 

s_pattern$month<-as.Date.yearmon(s_pattern$month)
s_pattern
```

#sector time series
```{r}
s_pattern %>% 
ggplot(aes(x=month, y=total_hrs))+
  geom_line(aes(color=Sector),size=1)+
  geom_point(aes(color=Sector),size=1.6)+
  scale_color_manual(values = c("#008253","#393d83","#fd7e14"))+
 # scale_x_yearmon(n=14)+
  scale_x_date(breaks = "1 month",date_labels="%b")+
  labs(
    y="Total hours",x=""
  )+
  geom_text(x =as.Date("2021-12-01"), y=600,label="Energy",
            color = "#393d83",family = f1,size = 2.2,face = "bold")+
  geom_text(x =as.Date("2021-12-01"), y=250,label="Climate change",
            color = "#008253",family = f1,size = 2.2,face = "bold")+
   geom_text(x =as.Date("2021-12-01"), y=150,label="Water",
            color = "#fd7e14",family = f1,size = 2.2,face = "bold")+
  ubuntu_theme()+
  theme(
    panel.grid.major.y = element_line(color = "grey78"),
    panel.grid.major.x = element_line(color = "grey78"),
    axis.text.x = element_text(angle = 70),
    axis.line = element_line(color = "white"),
    legend.position = "None"
  )
ggsave("./www/sector_series.png")
```

#plot
```{r}
p_pattern %>% 
ggplot(aes(x=month, y=total_hrs))+
  geom_line(aes(color=Lines),size=1)+
  geom_point(aes(color=Lines),size=1.6)+
  scale_color_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
 # scale_x_yearmon(n=14)+
  scale_x_date(breaks = "1 month",date_labels="%b")+
  labs(
    y="Total hours",x=""
  )+
   geom_text(x =as.Date("2021-12-01"), y=1200,label="Consulting",
            color = "#fd7e14",family = f1,size = 3,face = "bold")+
  geom_text(x =as.Date("2021-12-01"), y=500,label="Leave",
            color = "#ffc107",family = f1,size = 3,face = "bold")+
  geom_text(x =as.Date("2021-12-01"), y=150,label="BD",
            color = "#393d83",family = f1,size = 3,face = "bold")+
   geom_text(x =as.Date("2021-12-01"), y=300,label="Admin",
            color = "#008253",family = f1,size = 3,face = "bold")+
  ubuntu_theme()+
  theme(
    panel.grid.major.y = element_line(color = "grey78"),
    panel.grid.major.x = element_line(color = "grey78"),
    axis.text.x = element_text(angle = 70),
    axis.line = element_line(color = "white"),
    legend.position = "None"
  )
ggsave("./www/lines_series.png")
```

   #######Quarter 4 analysis############
```{r}
df_q4<-df %>% 
  filter(Quarter=="4")
  #select(-month2)
df_q4
```

#quarter 4 projects table
```{r}
sct_q4<-df_q4 %>% 
  filter(Lines=="Consulting") %>%
  group_by(Project,Sector) %>%
  summarise(total_cost=round(sum(Cost),0),
            total_hrs=round(sum(Hrs),0)) %>%
  arrange(desc(total_cost)) 
sct_q4
write.csv(sct_q4,"./www/q4_table2.csv",row.names = FALSE)
sct_q4b<-read_csv("./www/q4_table2.csv")
```

#table
```{r}

q4tb<-sct_q4b %>%
  gt() %>%
  gt_color_rows('total_cost':'total_hrs',palette = "ggsci::blue_material" ) %>%
  fmt_currency(
    columns = total_cost,
    currency = "USD",
    sep_mark = ",",
    decimals = NULL
  ) %>%
  cols_label(
    total_cost='Total cost ($)',
    total_hrs='Total hours'
  ) %>%
   tab_style(
    #select object to modify
    locations = cells_title(groups ="cell"),
    #specifying text style
    style = list(
      cell_text(
        size = "xx-large"
      )
    )
  ) %>%
  gt_theme_guardian()
#q4tb
q4tb %>% gtsave("./www/q4tb2.png")
```

#q4 lines
```{r}
lines_q4<-df_q4 %>% 
  group_by(Lines) %>%
  summarise(total_hrs=round(sum(Hrs),0),
            total_cost=sum(Cost)
           ) %>%
  mutate(
     hrs_perc=round(total_hrs/sum(total_hrs)*100,0)
  )


lines_q4[2,1]<-"BD"
lines_q4
```

```{r}
# Compute percentages
lines_q4$fraction <- lines_q4$hrs_perc / sum(lines_q4$hrs_perc)

# Compute the cumulative percentages (top of each rectangle)
lines_q4$ymax <- cumsum(lines_q4$fraction)

# Compute the bottom of each rectangle
lines_q4$ymin <- c(0, head(lines_q4$ymax, n=-1))

# Compute label position
lines_q4$labelPosition <- (lines_q4$ymax + lines_q4$ymin) / 2

# Compute a good label
lines_q4$label <- paste0(lines_q4$Lines, "\n", lines_q4$hrs_perc,"%")

##ggplot
p3<-ggplot(lines_q4, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=Lines)) +
  geom_rect() +
  geom_text( x=4.6, aes(y=labelPosition, label=label,color=Lines), size=3.2,family=f1) +
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107")) +
  scale_color_manual(values = c("#008253","#393d83","#fd7e14","#ffc107")) +
  coord_polar(theta="y") +
  xlim(c(1, 4)) +
  theme_void() +
  theme(legend.position = "none")
ggsave("./www/lines_q4.png",
       dpi = 320,
      width = 10,
      height = 7)
```


```{r}
sector_q4<-df_q4 %>% 
  #filter(Lines=="Consulting")%>%
  group_by(Sector) %>%
  summarise(total_hrs=round(sum(Hrs),0),
            total_cost=sum(Cost)
           ) %>%
  mutate(
     hrs_perc=round(total_hrs/sum(total_hrs)*100,1)
  )

sct<-c("Climate Change","Energy","Water")

sector_q4<-sector_q4 %>% filter(Sector %in% sct)
sector_q4

```
```{r}
p4<-ggplot(sector_q4,aes(x="",y=hrs_perc,fill=Sector))+
  geom_bar(position = "stack",stat = "identity",width = 0.2)+
  geom_text(aes(label=paste0(Sector,"\n",hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=2.5)+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14"))+
  # geom_curve(aes(x = 40, y = 3, xend = 0, yend = -6),
  #            size = 0.5, color = "black",curvature = -0.2,
  #            arrow = arrow(length = unit(0.02, "npc")))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    axis.title = element_blank(),
    axis.text.x = element_blank(),
    legend.position = "None",
    legend.title = element_blank()
  )
p4
```

```{r}
p6<-(p3|p4)
ggsave("./www/lines_q4.png",plot = p6, dpi = 320, height = 8, width = 11)
```

#time spent per line for each team q4
```{r}
team_linesq4<-df_q4 %>% 
  group_by(Team,Lines) %>%
  summarise(total_hrs=round(sum(Hrs),0),
            total_cost=sum(Cost)
           ) %>%
  mutate(
     hrs_perc=round(total_hrs/sum(total_hrs)*100,1)
  )
team_linesq4
```
```{r}
ggplot(team_linesq4,aes(x=Team,y=hrs_perc,fill=Lines))+
  geom_bar(position = "stack",stat = "identity")+
  geom_text(aes(label=paste0(hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=3)+
  labs(y="Percentage")+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )
ggsave("./www/team_linesq4.png")
```

#time spent for each line individually
```{r}
personal_linesq4<-df_q4 %>% 
  group_by(Name,Lines) %>%
  summarise(total_hrs=round(sum(Hrs),0),
            total_cost=sum(Cost)
           ) %>%
  mutate(
     hrs_perc=round(total_hrs/sum(total_hrs)*100,1)
  ) %>% 
  select(-c(3,4))

personal_linesq4a<-personal_linesq4 %>% spread(Lines,hrs_perc)


write.csv(personal_linesq4a,"./www/lines_tableq4.csv",row.names = FALSE)
personal_linesq4b<-read_csv("./www/lines_tableq4.csv")
```

#plotting table
```{r}
plq4tb<-personal_linesq4b %>%
  gt() %>%
   fmt_percent(
    columns = 'Admin':'Leave',
    scale_values = FALSE,
    decimals = 1
  )%>%
  gt_color_rows('Admin':'Leave',palette = "ggsci::blue_material" ) %>%
  
  gt_theme_guardian()

plq4tb %>% gtsave("./www/lines_tableq4b.png")
```
#time per person q4
```{r}
 psn_q4<-personal_linesq4 %>% mutate(
    Team=case_when(Name=="Barasa Murefu"~"Team Delta",
                 Name=="Kinuthia Brian"~"Team Phi",
                 Name=="Shikorire Tom"~"Team Phi",
                 Name=="Anyango Ruth"~"Team Phi",
                 Name=="Mwangi Brian"~"Team Phi",
                 Name=="Mbutura Alois"~"Team Kappa",
                 Name=="Kitetu Martin"~"Team Kappa",
                 Name=="Okode Adams"~"Team Kappa",
                 Name=="Irungu Joyce"~"Team Chi",
                 Name=="Odhiambo Elmah"~"Team Chi",
                 Name=="Samo June"~"Team Chi",
                 Name=="Cherunya Pauline"~"Team Chi",
                 Name=="Gichuhi Ruth"~"Team Gamma",
                 Name=="Muchiri James"~"Team Gamma",
                 Name=="Maina Morris"~"Team Gamma",
                 Name=="Muthuri Cynthia"~"Team Gamma"
                 )
  )
psn_q4
```
```{r}
psn_q4 %>% 
  filter(Team=="Team Phi")%>%
  ggplot(aes(x=Name,y=hrs_perc,fill=Lines))+
  geom_bar(position = "stack",stat = "identity")+
  geom_text(aes(label=paste0(hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=3)+
  labs(title ="Team Phi",y="Percentage",x="")+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )
ggsave("./www/tphi_q4.png")
```

```{r}
psn_q4 %>% 
  filter(Team=="Team Chi")%>%
  ggplot(aes(x=Name,y=hrs_perc,fill=Lines))+
  geom_bar(position = "stack",stat = "identity")+
  geom_text(aes(label=paste0(hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=3)+
  labs(title ="Team Chi",y="Percentage",x="")+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )
ggsave("./www/tchi_q4.png")
```

```{r}
psn_q4 %>% 
  filter(Team=="Team Gamma")%>%
  ggplot(aes(x=Name,y=hrs_perc,fill=Lines))+
  geom_bar(position = "stack",stat = "identity")+
  geom_text(aes(label=paste0(hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=3)+
  labs(title ="Team Gamma",y="Percentage",x="")+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )
ggsave("./www/tgamma_q4.png")
```

```{r}
psn_q4 %>% 
  filter(Team=="Team Delta")%>%
  ggplot(aes(x=Name,y=hrs_perc,fill=Lines))+
  geom_bar(position = "stack",stat = "identity")+
  geom_text(aes(label=paste0(hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=3)+
  labs(title ="Team Delta",y="Percentage",x="")+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )
ggsave("./www/tdelta_q4.png")
```

```{r}
psn_q4 %>% 
  filter(Team=="Team Kappa")%>%
  ggplot(aes(x=Name,y=hrs_perc,fill=Lines))+
  geom_bar(position = "stack",stat = "identity")+
  geom_text(aes(label=paste0(hrs_perc,"%")),position = position_stack(vjust = 0.5),color="white",family=f1,size=3)+
  labs(title ="Team Kappa",y="Percentage",x="")+
  guides(fill=guide_legend(reverse = TRUE))+
  scale_fill_manual(values = c("#008253","#393d83","#fd7e14","#ffc107"))+
  coord_flip()+
  ubuntu_theme()+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )
ggsave("./www/tkappa_q4.png")
```

#time spent on each project
```{r}
tprojectq4<- df_q4 %>% 
  filter(Lines=="Consulting") %>%
  group_by(Project,Team) %>%
  summarise(total_hrs=round(sum(Hrs),0),
            total_cost=sum(Cost)
           ) 
  
tprojectq4
```
```{r}
tprojectq4 %>% 
  filter(Team=="Team Gamma") %>%
  ggplot(aes(x=reorder(Project,total_hrs),y=total_hrs))+
  geom_col(fill="#008253")+
  labs(title ="Team Gamma",y="Total hours",x="")+
  geom_text(aes(label=paste0(total_hrs)),position=position_dodge(width=1.2),color="black",
            hjust=0,vjust=0.5,family=f1,size=2.8)+
  coord_flip()+
  ubuntu_theme()
ggsave("./www/tprojectq4_gamma.png")
```


```{r}
tprojectq4 %>% 
  filter(Team=="Team Phi") %>%
  ggplot(aes(x=reorder(Project,total_hrs),y=total_hrs))+
  geom_col(fill="#393d83")+
  labs(title ="Team Phi",y="Total hours",x="")+
  geom_text(aes(label=paste0(total_hrs)),position=position_dodge(width=1.2),color="black",
            hjust=0,vjust=0.5,family=f1,size=2.6)+
  coord_flip()+
  ubuntu_theme()
ggsave("./www/tprojectq4_phi.png")
```

```{r}
tprojectq4 %>% 
  filter(Team=="Team Chi") %>%
  ggplot(aes(x=reorder(Project,total_hrs),y=total_hrs))+
  geom_col(fill="#fd7e14")+
  labs(title ="Team Chi",y="Total hours",x="")+
  geom_text(aes(label=paste0(total_hrs)),position=position_dodge(width=1.2),color="black",
            hjust=0.6,vjust=0.5,family=f1,size=2.4)+
  coord_flip()+
  ubuntu_theme()
ggsave("./www/tprojectq4_chi.png")
```

```{r}
tprojectq4 %>% 
  filter(Team=="Team Kappa") %>%
  ggplot(aes(x=reorder(Project,total_hrs),y=total_hrs))+
  geom_col(fill="#ffc107")+
  labs(title ="Team Kappa",y="Total hours",x="")+
  geom_text(aes(label=paste0(total_hrs)),position=position_dodge(width=1.2),color="black",
            hjust=0.6,vjust=0.5,family=f1,size=2.4)+
  coord_flip()+
  ubuntu_theme()
ggsave("./www/tprojectq4_kappa.png")
```

```{r}
tprojectq4 %>% 
  filter(Team=="Team Delta") %>%
  ggplot(aes(x=reorder(Project,total_hrs),y=total_hrs))+
  geom_col(fill="#2596be")+
  labs(title ="Team Delta",y="Total hours",x="")+
  geom_text(aes(label=paste0(total_hrs)),position=position_dodge(width=1.2),color="black",
            hjust=0,vjust=0.5,family=f1,size=2.4)+
  coord_flip()+
  ubuntu_theme()
ggsave("./www/tprojectq4_delta.png")
```



